Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database

### 2020-09-25 manuscript/ERL_global_C_review.Rmd
update and add a couple summary tables #29

```{r eval = TRUE, echo=FALSE, warning=FALSE}
#READING IN TABLES TO GET STATS FOR TEXT

#ForC raw
ForC_measurements <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_measurements.csv", stringsAsFactors = FALSE, check.names = FALSE)
  n_records_age_unk <- sum(is.na(as.numeric(ForC_measurements[, "stand.age"])))  #total n records no age info
  n_SRDB=NA
  n_GROA=NA
ForC_plots <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_plots.csv", stringsAsFactors = FALSE, check.names = FALSE)
ForC_sites <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_sites.csv", stringsAsFactors = FALSE, check.names = FALSE)
  n_sites_coord_unk <- sum(is.na(as.numeric(ForC_sites[, "lat"])))  #total n records no geographic coordinates
  n_areas <- max(ForC_sites[,"geographic.area"], na.rm = TRUE )

#ForC simplified
ForC_simplified <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/ForC_simplified/ForC_simplified.csv", stringsAsFactors = FALSE, check.names = FALSE)
n_no_history <- sum(ForC_simplified[,"history.no.info"]) #(relative to ForC_simplified)
n_managed_in_ForC_simplified <- sum(ForC_simplified[,"managed"])
n_disturbed_in_ForC_simplified <- sum(ForC_simplified[,"disturbed"])
records_eligible_for_inclusion <- NA  #number in ForC_simplified that are not disturbed, managed, or no.history

# limiting to variables and biomes of interest:
n_focal_variable_records_eligible_for_inclusion <- NA # records_eligible_for_inclusion, limited to focal variables

@@

n_plots_tot <- NA #total numbers of plots used in the analysis (all variables combined)
n_areas_tot <- NA #total numbers of areas used in the analysis (all variables combined)

#numbers of variables with records from ≥7 geographic areas, by biomes/ age class:
n_fluxes_matureTrB_with_ge7_areas <- NA
n_stocks_matureTrB_with_ge7_areas <- NA
n_fluxes_matureTeB_with_ge7_areas <- NA
n_stocks_matureTeB_with_ge7_areas <- NA
n_fluxes_matureTeN_with_ge7_areas <- NA
n_stocks_matureTeN_with_ge7_areas <- NA
n_fluxes_matureBoN_with_ge7_areas <- NA
n_stocks_matureBoN_with_ge7_areas <- NA

n_fluxes_youngTrB_with_ge7_areas <- NA
n_stocks_youngTrB_with_ge7_areas <- NA
n_fluxes_youngTeB_with_ge7_areas <- NA
n_stocks_youngTeB_with_ge7_areas <- NA
n_fluxes_youngTeN_with_ge7_areas <- NA
n_stocks_youngTeN_with_ge7_areas <- NA
n_fluxes_youngBoN_with_ge7_areas <- NA
n_stocks_youngBoN_with_ge7_areas <- NA

@@

For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature). 
Climate zones (Fig. 1) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd). 
Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis. 
Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006]. 
Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication). 
Assigning stands to these groupings required the exclusion of records for which ForC lacked records of stand age (n=`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database), or geographic coordinates (n=`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database). 
In total, our analysis retained `r round(n_records/n_focal_variable_records_eligible_for_inclusion*100, digits=1)` of the focal variable records for forests of known age.

@@

Of the `r nrow(ForC_measurements)` records in ForC v.#, `r n_records` met the criteria for inclusion in this study (Fig. 1). 
These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas. 
Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained mature forest estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_matureTrB_with_ge7_areas` fluxes and `r n_stocks_matureTrB_with_ge7_areas` stocks in tropical broadleaf forests, `r n_fluxes_matureTeB_with_ge7_areas` fluxes and `r n_stocks_matureTeB_with_ge7_areas` stocks in temperate broadleaf forests, `r n_fluxes_matureTeN_with_ge7_areas` fluxes and `r n_stocks_matureTeN_with_ge7_areas` stocks in temperate conifer forests, and `r n_fluxes_matureBoN_with_ge7_areas` fluxes and `r n_stocks_matureBoN_with_ge7_areas` stocks in boreal forests. 
For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_youngTrB_with_ge7_areas` fluxes and `r n_stocks_youngTrB_with_ge7_areas` stocks in tropical broadleaf forests, `r n_fluxes_youngTeB_with_ge7_areas` fluxes and `r n_stocks_youngTeB_with_ge7_areas` stocks in temperate broadleaf forests, `r n_fluxes_youngTeN_with_ge7_areas` fluxes and `r n_stocks_youngTeN_with_ge7_areas` stocks in temperate conifer forests, and `r n_fluxes_youngBoN_with_ge7_areas` fluxes and `r n_stocks_youngBoN_with_ge7_areas` stocks in boreal forests.

@@

  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Tropical broadleaf MATURE.png")
```

@@

  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate broadleaf MATURE.png")
```

@@

  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate conifer MATURE.png")
```

@@

  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Boreal conifer MATURE.png")
```
\elandscape

@@

  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Tropical broadleaf YOUNG.png")
```

@@

  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate broadleaf YOUNG.png")
```

@@

  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Temperate conifer YOUNG.png")
```

@@

  knitr::include_graphics("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/figures/C_cycle_diagrams/Diagrams/Boreal conifer YOUNG.png")
```
\elandscape

>>>>

```{r eval = TRUE, echo=FALSE, warning=FALSE}
#READING IN TABLES TO GET STATS FOR TEXT
path_to_ForC <- paste0(dirname(dirname(getwd())), "/ForC/") # this is to allow anybody to knit (assuming current working direcoty is in forc-db/ERL-review/manuscript and its "nesting" withing the ForC repo does not change)

#ForC raw
ForC_measurements <- read.csv(paste0(path_to_ForC, "/data/ForC_measurements.csv"), stringsAsFactors = FALSE, check.names = FALSE)
  n_records_age_unk <- sum(is.na(as.numeric(ForC_measurements[, "stand.age"])))  #total n records no age info
  n_records_age_0 <- sum(as.numeric(ForC_measurements[, "stand.age"]) %in% 0)

  n_SRDB <- sum(ForC_measurements$ForC.investigator %in% "Ben Bond-Lamberty")
  n_GROA <- sum(ForC_measurements$ForC.investigator %in% "Dr. Susan Cook-Patton")

  ForC_plots <- read.csv(paste0(path_to_ForC, "/data/ForC_plots.csv"), stringsAsFactors = FALSE, check.names = FALSE)
ForC_sites <- read.csv(paste0(path_to_ForC, "/data/ForC_sites.csv"), stringsAsFactors = FALSE, check.names = FALSE)

n_sites_coord_unk <- sum(is.na(as.numeric(ForC_sites[, "lat"])))  #total n records no geographic coordinates

n_areas <- length(unique(na.omit(ForC_sites[,"geographic.area"])))

#ForC simplified
ForC_simplified <- read.csv(paste0(path_to_ForC, "ForC_simplified/ForC_simplified.csv"), stringsAsFactors = FALSE, check.names = FALSE)

n_no_history <- sum(ForC_simplified[,"history.no.info"]) #(relative to ForC_simplified)
n_managed_in_ForC_simplified <- sum(ForC_simplified[,"managed"])
n_disturbed_in_ForC_simplified <- sum(ForC_simplified[,"disturbed"])
records_eligible_for_inclusion <- sum(ForC_simplified$managed ==0 & ForC_simplified$disturbed ==0 & ForC_simplified$history.no.info == 0)  #number in ForC_simplified that are not disturbed, managed, or no.history

# limiting to variables and biomes of interest:
 Biomes_n_records <- read.csv("tables_figures/Biomes_n_records_variables_of_interest.csv", check.names = F)
n_focal_variable_records_eligible_for_inclusion <- sum(Biomes_n_records$n_records) # n records_eligible_for_inclusion, limited to focal variables

@@

global_counts <- read.csv("tables_figures/Count_n_record_n_plot_n_area.csv", check.names = FALSE)
n_records - global_counts$n_records # one count difference... not sure why. probably the filter thing... will ignore dor now.
n_plots_tot <- global_counts$n_plots #total numbers of plots used in the analysis (all variables combined)
n_areas_tot <- global_counts$n_areas #total numbers of areas used in the analysis (all variables combined)

#numbers of variables with records from ≥7 geographic areas, by biomes/ age class:
ForC_biome_averages <- read.csv(paste0(path_to_ForC, "numbers_and_facts/ForC_variable_averages_per_Biome.csv"), stringsAsFactors = F)

idx_fluxes <- ForC_biome_averages$n.areas >= 7 & ForC_biome_averages$variable.type %in% "flux"
idx_stocks <- ForC_biome_averages$n.areas >= 7 & ForC_biome_averages$variable.type %in% "stock"

n_fluxes_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_fluxes])
n_stocks_with_ge7_areas <- table(ForC_biome_averages$Biome[idx_stocks])

@@

For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature). 
Climate zones (Fig. 1) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd). 
Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis. 
Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006]. 
Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication). 
Assigning stands to these groupings required the exclusion of records for which ForC lacked records of stand age (n=`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database), or geographic coordinates (n=`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database). We also excluded records of stand age = 0 year (n=`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
In total, our analysis retained `r round(n_records/n_focal_variable_records_eligible_for_inclusion*100, digits=1)` of the focal variable records for forests of known age.

@@

Of the `r nrow(ForC_measurements)` records in ForC v.#, `r n_records` met the criteria for inclusion in this study (Fig. 1). 
These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas. 
Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained mature forest estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests. 
For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_with_ge7_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_fluxes_with_ge7_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.

@@

  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Tropical broadleaf MATURE.png"))
```

@@

  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate broadleaf MATURE.png"))
```

@@

  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate conifer MATURE.png"))
```

@@

  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Boreal conifer MATURE.png"))
```
\elandscape

@@

  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Tropical broadleaf YOUNG.png"))
```

@@

  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate broadleaf YOUNG.png"))
```

@@

  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Temperate conifer YOUNG.png"))
```

@@

  knitr::include_graphics(paste0(path_to_ForC, "figures/C_cycle_diagrams/Diagrams/Boreal conifer YOUNG.png"))
```
\elandscape
